Global patterns of forest autotrophic carbon fluxes

### 2020-08-07 manuscript/manuscript draft.Rmd
general clean-up

---
title: ""
author: ""
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
bibliography: [library.bib,  packages.bib]
biblio-style: apalike
csl: global-change-biology.csl
header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}
                  \usepackage{setspace}\onehalfspacing
                  \usepackage{lineno}\linenumbers

@@

2. *Becky- current*

@@

**Keywords:**

**Paper type:** Primary Research Article

\newpage
### Abstract
Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
Here, we draw upon 1319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
Temperature variables in general, and mean annual temperature ($MAT$) and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed. 
The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
C fluxes increased with growing season length, but this was never the best univariate predictor.
Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
These findings clarify how forest C flux varies with latitude and climate on a global scale. 
In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.

@@

Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008]. 
Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018]. 
They are linked to cycling of energy, water, and nutrients, and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019]-- a flux more than seven times greater that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019]. 
As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and/or management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].

Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019], but studies have differed in their conclusions regarding the shape of this relationship - quite possibly because of lack of standardization with respect to methodology and stand history.
For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. 
In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007]--or even a decrease [@huston_global_2009, but see @gillman_latitude_2015]. 
A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
Such differences among C fluxes their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity. However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].

The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010].
However, there is little consensus as to the shapes of these relationships or the best predictor variables.
The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables. 
C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @wei_forest_2010] with increasing $MAT$.
In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
C fluxes generally saturate at high levels of MAP, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003]. 
Interactions between MAT and MAP are also possible; within the tropics, there is a positive interaction between MAT and MAP in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017]. 
There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.

As metrics of annual climate, MAT and MAP fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
Some studies have suggested that the apparently strong relationship between MAT and C fluxes is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018]. 
@kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to growing season length [but see @chu_does_2016].
While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, analysis with carefully standardized variables and forest ages would be necessary to test the the veracity and generality of this hypothesis.

The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude.
We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on C flux standardized by growing season length.

@@

This analysis focused on nine C flux variables included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010]. Additional targeted literature searches were conducted to identify further available data on the fluxes analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016. We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.

This analysis focused on mature forests with no known history of signficant disturbance or management. 
There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015]. To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth". 
Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. 
Sites that had experienced significant disturbance within the past 100 years were also excluded. Disturbances that qualified sites for exclusion included major cutting or harvesting, burning, flooding, drought and storm events with site mortality >10% of trees. Grazed sites were retained.

@@

*Climate data*

ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature (MAT) and mean annual precipitation (MAP) as reported in the primary literature [@anderson-teixeira_forc_2018]. Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity (MAP/PET), and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1). From these data, we derived maximum VPD, defined as the VPD of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than PET. Where site-level data was missing for MAT or MAP, we used values from the WorldClim dataset.

Following the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014], length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C. 
We experimented with a definition of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here, and so chose to exclude it. 
Monthly data for PET, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season. 
Total growing season precipitation and solar radiation were also calculated.

@@

The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. 
The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. 
Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. 
We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$ AIC) relative to models excluding altitude was generally small (often $\Delta$ AIC<2). 
Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta$ AIC $\ge$ 2.0 relative to a corresponding null model. 
All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
Specific analyses are as described below.

We first examined the relationship between latitude and C fluxes (*Q1*; Table 1). 
We tested models with latitude as a first-order linear, second-order polynomial, and logarithmic term. 
For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.

@@

Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$. For each group, measurements taken at the same site and plot, an in the same year were grouped together. For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third. The ratio of each pair of component fluxes was calculated. The log of these ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and extreme outliers removed,

We next examined the relationships of C fluxes to climate variables (*Q2-Q4*; Table 1). 
We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta$ AIC $\ge$ 2.0 relative to a first-order linear model.
We tested relationships of each C flux (Table 2) against each climate variable (Table S1). Variables which were not significant explanatory variables or which explained <20% of variation in C fluxes are only presented in SI.

Multivariate models were used to investigate the potential joint and interactive effects of MAT and MAP on carbon fluxes. An additive model including MAP in addition to MAT was accepted when $\Delta$ AIC >2 relative to a null including only MAT as a fixed effect. An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.

@@

**All analyses were conducted in R (**Version**). Code and data necessary to reproduce all results are archived on GitHub....**

@@

In total, we analyzed 1319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig. 1, Table 2).

*How does C flux vary with latitude?*

All major carbon fluxes decreased with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001). The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18). The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly - but not significantly - better fit.

@@

In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig. 2). That is, modeled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.

We found no evidence of systematic variation in C allocation with latitude or climate (Fig. S3). Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, MAT, MAP and temperature seasonality), none were significant.

*How does C flux relate to MAT and MAP?*

All fluxes increased with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
For eight of the nine fluxes, this relationship was linear.
For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
*As with latitude, MAT tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**

>>>>

---
title: ""
author: ""
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
bibliography: [library.bib,  packages.bib]
biblio-style: apalike
csl: global-change-biology.csl
header-includes:  \usepackage{float}
                  \usepackage{caption}
                  \captionsetup[table]{font=footnotesize}
                  \captionsetup[figure]{font=footnotesize}
                  \usepackage{setspace}\doublespacing
                  \usepackage{lineno}\linenumbers

@@

2. **Becky- current**

@@

\newpage
### Abstract
Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale. 
We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests. 
C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
Temperature variables in general, and mean annual temperature ($MAT$) and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed. 
The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
C fluxes increased with growing season length, but this was never the best univariate predictor.
Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
These findings clarify how forest C flux varies with latitude and climate on a global scale. 
In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.

@@

Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [Millenium Ecosystem @millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008]. 
Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018]. 
They are linked to cycling of energy, water, and nutrients and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019]-- a flux more than seven times greater that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019]. 
As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].

Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019]. 
However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history.
For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. 
In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007] -- or even a decrease [@huston_global_2009, but see @gillman_latitude_2015]. 
A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
Such differences among C fluxes their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity. However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].

The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010].
However, there is little consensus as to the shapes of these relationships or the best predictor variables.
The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature ($MAT$) and mean annual precipitation ($MAP$), as the most commonly reported site-level climate variables. 
C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @wei_forest_2010] with increasing $MAT$.
In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003]. 
Interactions between $MAT$ and $MAP$ are also possible; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017]. 
There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.

As metrics of annual climate, $MAT$ and $MAP$ fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
Some studies have suggested that the apparently strong relationship between $MAT$ and C fluxes is actually a consequence of the correlation between $MAT$ and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018]. 
@kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to growing season length [but see @chu_does_2016].
While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, analysis with carefully standardized variables and forest ages would be necessary to test the the veracity and generality of this hypothesis.

The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude (*Q1*).
We then test how these fluxes relate to $MAT$ and $MAP$ (*Q2*), and additionally how they respond to other, less well-studied, climate variables (*Q3*).
Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes (*Q4*), and the influence of climate on C flux standardized by growing season length (*Q5*).

@@

This analysis focused on nine C flux variables included in the open-access ForC database [Table 2; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. 
ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. 
Associated data, such as stand age, measurement methodologies, and disturbance history, are also included. The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010]. 
Additional targeted literature searches were conducted to identify further available data on the fluxes analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016. 
We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. 
This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
From this, we drew 1,319 records that met our criteria, as outlined below (Fig. 1).

This analysis focused on mature forests with no known history of significant disturbance or management. 
There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015]. To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth". 
Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description. 
Sites that had experienced significant disturbance within the past 100 years were also excluded. Disturbances that qualified sites for exclusion included major cutting or harvesting, burning, flooding, drought and storm events with site mortality >10% of trees. Grazed sites were retained.

@@

\newpage  
*Climate data*

ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018]. Based on the geographic co-ordinates for each site, data on twelve climate variables--including $MAT$, $MAP$, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1). From these data, we derived maximum $VPD$, defined as the $VPD$ of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than $PET$. Where site-level data was missing for $MAT$ or $MAP$, we used values from the WorldClim dataset.

Following the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014], length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C. 
We experimented with a definition of growing season months including a moisture index, defined as $(MAT - PET)/PET > -0.95$ [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here, and so chose to exclude it. 
Monthly data for $PET$, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly $PET$, precipitation, temperature and solar radiation during the growing season. 
Total growing season precipitation and solar radiation were also calculated.

@@

The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. 
The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. 
Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018]. 
We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2). 
Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
Specific analyses are as described below.

We first examined the relationship between latitude and C fluxes (*Q1*; Table 1). 
We tested models with latitude as a first-order linear, second-order polynomial, and logarithmic term. 
For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.

@@

Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$. For each group, measurements taken at the same site and plot, an in the same year were grouped together. For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third. The ratio of each pair of component fluxes was calculated. The log of these ratios were regressed against latitude and climate variables, using the linear model specified above. Cook's distance analyses were carried out for each of the models, and extreme outliers removed.

We next examined the relationships of C fluxes to climate variables (*Q2-Q4*; Table 1). 
We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta AIC$ $\ge$ 2.0 relative to a first-order linear model.
We tested relationships of each C flux (Table 2) against each climate variable (Table S1). Variables which were not significant explanatory variables or which explained <20% of variation in C fluxes are only presented in SI.

Multivariate models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes. An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta AIC$ >2 relative to a null including only $MAT$ as a fixed effect. An interactive model including an $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.

@@

All analyses were conducted in R (**Version**). Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).

@@

In total, we analyzed 1,319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years. These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Figs. 1, S1;  Table 2).

*Q1. How does C flux vary with latitude?*

All major carbon fluxes decreased with latitude (Fig. 2; Table S2). Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2). Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001). The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18). The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly -- but not significantly -- better fit.

@@

Smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig. 2). That is, modeled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.

We found no evidence of systematic variation in C allocation with latitude or climate (Fig. S3). Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.

*Q2. How does C flux relate to $MAT$ and MAP?*

All fluxes increased with $MAT$ (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
For eight of the nine fluxes, this relationship was linear.
**For only one variable, $BNPP$, did a lognormal fit provide significant improvement over a first-order linear relationship.** 
As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).

$MAP$ was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
However, it explained little variation: with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux. All fluxes increased with $MAP$ up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).
